---
title: "Homework 3"
author: "Group B: Abdalghani, Demirbilek, Plasencia Palacios, Spagnolo"
date: "Spring 2020"
output:
  html_document:
    toc: no
header-includes:
- \usepackage{color}
- \definecolor{Purple}{HTML}{911146}
- \definecolor{Orange}{HTML}{CF4A30}
- \setbeamercolor{alerted text}{fg=Orange}
- \setbeamercolor{frametitle}{bg=Purple}
institute: University of Udine & University of Trieste
graphics: yes
fontsize: 10pt
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.align = 'center', warning=FALSE, message=FALSE, fig.asp=0.625, dev='png', global.par = TRUE, dev.args=list(pointsize=10), fig.path = 'figs/', fig.height = 10, fig.width = 10)
```

```{r setup, include=FALSE}
library(MASS)
library(knitr)
local({
  hook_plot = knit_hooks$get('plot')
  knit_hooks$set(plot = function(x, options) {
    paste0('\n\n----\n\n', hook_plot(x, options))
  })
})
```

# {.tabset}

## DAAG {.tabset}

### Chapter 6 - Exercise 10


The data frame `table.b3` in the $MPV$ package contains data on gas mileage and 11 other variables for a sample of 32 automobiles.

a) Construct a scatterplot of `y` (mpg) versus `x1` (displacement). Is the relationship between these variables non-linear?

```{r echo=TRUE,  message=FALSE, warning=FALSE}
library(MPV)
automobiles <- table.b3

#help(table.b3)

plot(y~x1, data = automobiles, main = "y vs x1 (displacement)",xlab = "x1 (displacement)", ylab = "y", pch = 19,col="purple")

```
The scatterplot shows curvilinear relationship between y and x1.

b) Use the `xyplot()` function, and `x11` (type of transmission) as a `group` variable. Is a linear model reasonable for these data?

```{r echo=TRUE,  message=FALSE, warning=FALSE}
library(lattice)
xyplot(y ~ x1 + x2 + x3, data = automobiles, groups = x11, main = "y vs x", pch = 19, key=list(rep=FALSE,columns=1,
                                                                                          text=list(lab=c("manual")),pch=19,points=list(col="cadetblue3"),
                                                                                          text=list(lab=c("automatic")),points=list(col="purple")))

```
This plot shows that apparent nonlinearity can be explained by the two types of transmission.

c) Fit the model relating `y` to `x1` and `x11` which gives two lines having possibly different slopes and intercepts. Check the diagnostics. Are there any influential observations? Are there any influential outliers?

```{r echo=TRUE,  message=FALSE, warning=FALSE}
model0 <- lm(y ~ x1*x11, data = automobiles)
par(mfrow = c(1,4))
plot(model0)

```
Residual plot show no systematic behaviour. Observation 5 ,15, 22 influential and 5 is possible outlier according to Cook's distance.


d) Plot the residuals against the variable `x7` (number of transmission speeds), again using `x11` as a group variable. Is there anything striking about this plot?

```{r echo=TRUE,  message=FALSE, warning=FALSE}
xyplot(resid(model0) ~ x7, data=automobiles, group=x11, main = "model0_residuals vs x7 (number of transmission speeds)", pch = 19,
       key=list(rep=FALSE,columns=1,text=list(lab=c("manual")),pch=19,points=list(col="cadetblue3"),text=list(lab=c("automatic")),points=list(col="purple")))

```
Here observation 5 is the only car which is manual and has 3-speed transmisson so we can say that our previous foresight was right, 5 is a strong candiate as an outlier.

### Chapter 8 - Exercise 3

Consider again the `moths` data set of Section 8.4.

a) What happens to the standard error estimates when the `poisson` family is used in `glm()` instead of the `quasipoisson` family?

```{r echo=TRUE,  message=FALSE, warning=FALSE}
library(DAAG)
#help(moths)
#moths

summary(A.glm <-  glm(A ~ habitat + log(meters), family = quasipoisson, data = moths)) # taken from DAAG book
summary(A.glm <-  glm(A ~ habitat + log(meters), family = poisson, data = moths))

```
We can see that dispersion parameter for quassipoisson is 2.7. One of the explanation to this might be there is a clustering effect with an average cluster size 2.7. Therefore standard errors and p-values from a model that assumed Poisson errors would be misleading. Using quassipoisson has increased the standard errors by a factor of $\sqrt{2.7}=1.64$. So using poisson will reduce the SE by this factor if the poisson family specified.

b) Analyze the `P` moths, in the same way as the A moths were analyzed. Comment on the effect of transect length.

```{r echo=TRUE,  message=FALSE, warning=FALSE}
sapply(split(moths$P, moths$habitat), sum)

dotplot(habitat ~ P ,data = moths,  xlab="Number of moths (species P)",
  panel=function(x, y, ...){
    panel.dotplot(x,y, pch=1,type="p", col="red", ...)
    panel.average(x, y, pch=3, cex=1.25, type="p", col="blue")
    },
  key=list(text=list(c("Individual transects", "Mean")),
    points=list(pch=c(1,3), cex=c(1,1.25), col=c("red","blue")),
    columns=2))

# didn't understand why blue is still line even though I specified it as pch=3
#?relevel

moths$habitat <- relevel(moths$habitat, ref = "Lowerside")
summary(P.glm <- glm(P ~ habitat + log(meters), family =  quasipoisson, data = moths)) 

```
The highest numbers are in habitats `SWsoak` and `Disturbed`. Number of moths increases with the transect lenght, by a factor of $e^{0.55} =1.73 $ for each one meter increase in transect lenght.