---
title: "Homework 3"
author: "Group B: Abdalghani, Demirbilek, Plasencia Palacios, Spagnolo"
date: "Spring 2020"
output:
  html_document:
    toc: no
header-includes:
- \usepackage{color}
- \definecolor{Purple}{HTML}{911146}
- \definecolor{Orange}{HTML}{CF4A30}
- \setbeamercolor{alerted text}{fg=Orange}
- \setbeamercolor{frametitle}{bg=Purple}
institute: University of Udine & University of Trieste
graphics: yes
fontsize: 10pt
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.align = 'center', warning=FALSE, message=FALSE, fig.asp=0.625, dev='png', global.par = TRUE, dev.args=list(pointsize=10), fig.path = 'figs/', fig.height = 10, fig.width = 10)
```

```{r setup, include=FALSE}
library(MASS)
library(knitr)
local({
  hook_plot = knit_hooks$get('plot')
  knit_hooks$set(plot = function(x, options) {
    paste0('\n\n----\n\n', hook_plot(x, options))
  })
})
```

# {.tabset}

## DAAG {.tabset}

### Chapter 6 - Exercise 10


The data frame `table.b3` in the $MPV$ package contains data on gas mileage and 11 other variables for a sample of 32 automobiles.

a) Construct a scatterplot of `y` (mpg) versus `x1` (displacement). Is the relationship between these variables non-linear?

b) Use the `xyplot()` function, and `x11` (type of transmission) as a `group` variable. Is a linear model reasonable for these data?

c) Fit the model relating `y` to `x1` and `x11` which gives two lines having possibly different slopes and intercepts. Check the diagnostics. Are there any influential observations? Are there any influential outliers?

d) Plot the residuals against the variable `x7` (number of transmission speeds), again using `x11` as a group variable. Is there anything striking about this plot?




```{r echo=TRUE,  message=FALSE, warning=FALSE}


```

### Chapter 8 - Exercise 3

Consider again the `moths` data set of Section 8.4.

a) What happens to the standard error estimates when the `poisson` family is used in `glm()` instead of the `quasipoisson` family?

b) Analyze the `P` moths, in the same way as the A moths were analyzed. Comment on the effect of transect length.


```{r echo=TRUE,  message=FALSE, warning=FALSE}


```